An Extension of LTL with Rules\\ and its Application to Runtime Verification

### 2019-02-03 main.tex
fixing problems

\title{An Extension of LTL with Rules 
and its Application to Runtime Verification}

@@

@@

Hence, the sequence of states 
\[ \{ r \} \{ v(1) \}  \{ v(2) \} \{ v(3) \}  \{ r\}  \{ v(2) \}  \{  v(1) \}  \{  v(3) \} \{  r \} \] is a model of this formula, while \[ \{ r \} \{ v(1) \}  \{ v(2) \}  \{ v(2) \}  \{ r\}  \{ v(2) \}  \{  v(1) \}  \{  v(5) \}  \{  r \}\] is not.
The formula is a bit involved, hence given in several parts (macros).
\begin{itemize}
\item $add(P) ::= \forall x \, (P(x) \leftrightarrow (v(x) \vee \ominus P(x))$ [The set $P$ differs in the current state from the previous one only by values $a$ that appear in $v(a)$.]
\item $same(P) ::= \forall x \, (P(x) \leftrightarrow \ominus P(x))$ 
[The set of elements $x$ for which
$P (x)$ holds
contains the same elements as in the previous state.]
\item $empty (P) ::= \forall x \neg P(x)$ [The set $P$ is empty.]
\item $equal (P,Q) ::= \forall x (P(x) \leftrightarrow Q(x))$ 
[The sets represented by the relations
$P$ and $Q$ contain the same elements {\em in the current state}.]
\item $collect(P, \varphi, \psi) ::=
       q \wedge same(P) \wedge \varphi \wedge 
       \ominus ((add(P)\wedge \neg q \wedge \varphi) \since (q \wedge empty(P) \wedge \psi))$  [Collect elements $a$ in $v(a)$, while $\varphi$ holds, 
       since the previous state where $q$ 
       and $\psi$ held. ]
\item $init ::= r \wedge \Box (r \leftrightarrow \neg \exists x \, v(x))$ [$r$ holds in the first state, and $v(a)$ with some value $a$ holds in states where $q$ does not hold.]
\item $init \wedge \neg \Exists A \Exists B (equal (A, B) \wedge \\
collect(B, same(A),\, same(A) \since collect(A, true, true))$
[There are no two sequences of states where $v(a)$ holds
with various values $a$, between a pair of states satisfying $r$ that contain exactly the same elements (regardless of repetition).]
\end{itemize}
We prove that the above property cannot be expressed in EFLTL.
We use a combinatorial argument.
In EFLTL, after every prefix of the input, it is
sufficient to keep a finite set of (non temporal) relations over the
data elements seen in the input, which
summarizes the prefix. This follows from the set
semantics, which appears in Section~\ref{EPLTLRV}, 
where the interpretation of
each subformula, and the auxiliary relations,
at a particular state is a relation, and the interpretation 
of a property a prefix of length $i$ only depends on
these relations. Now, the number of such possible relations
for a fixed number of data elements $N$ is polynomial
in $N$. However, the above property checks whether
a set of values that appear within the relation $v$, between two consecutive appearances
of $r$'s, are the same. For that, one needs to keep a set of
subsets of the elements that appeared in the input so far. 
But the number of such subsets is exponential in $N$. \qed

>>>>

\title{An Extension of LTL with Rules\\ 
and its Application to Runtime Verification}

@@

\klaus{In the text below mention what $i$ is (time point, position, ...}

@@

\klaus{Below: isn't it the other way around, the first is not a model and the second is a model?}

Hence, the sequence of states 
\[ \{ r \} \{ v(1) \}  \{ v(2) \} \{ v(3) \}  \{ r\}  \{ v(2) \}  \{  v(1) \}  \{  v(3) \} \{  r \} \] is a model of this formula, while \[ \{ r \} \{ v(1) \}  \{ v(2) \}  \{ v(2) \}  \{ r\}  \{ v(2) \}  \{  v(1) \}  \{  v(5) \}  \{  r \}\] is not.
The QFLTL formula is a bit involved, hence given in several parts (macros).

\begin{itemize}
\item $add(P) ::= \forall x \, (P(x) \leftrightarrow (v(x) \vee \ominus P(x))$ [The set $P$ differs in the current state from the previous one only by values $a$ that appear in $v(a)$.]
\item $same(P) ::= \forall x \, (P(x) \leftrightarrow \ominus P(x))$ 
[The set of elements $x$ for which
$P (x)$ holds
contains the same elements as in the previous state.]
\item $empty (P) ::= \forall x \neg P(x)$ [The set $P$ is empty.]
\item $equal (P,Q) ::= \forall x (P(x) \leftrightarrow Q(x))$ 
[The sets represented by the relations
$P$ and $Q$ contain the same elements {\em in the current state}.]
\item $collect(P, \varphi, \psi) ::=
       r \wedge same(P) \wedge \varphi \wedge 
       \ominus ((add(P)\wedge \neg r \wedge \varphi) \ \since \ (r \wedge empty(P) \wedge \psi))$  [Collect elements $a$ in $v(a)$, while $\varphi$ holds, 
       since the previous state where $r$ 
       and $\psi$ held.]
\item $init ::= r \wedge \Box (r \leftrightarrow \neg \exists x \, v(x))$ [$r$ holds in the first state, and $v(a)$ with some value $a$ holds in states where $r$ does not hold.]
\item $init \wedge \neg \Exists A \Exists B (equal (A, B) \wedge \\
collect(B, same(A),\, same(A)) \ \since \ collect(A, true, true))$
[There are no two sequences of states where $v(a)$ holds
with various values $a$, between a pair of states satisfying $r$ that contain exactly the same elements (regardless of repetition).]
\end{itemize}

\klaus{Above: maybe need to point out what the top formula is.}

\klaus{Above: isn't there missing a $\Box$ in the main formula?}

We prove that the above property cannot be expressed in EFLTL.
We use a combinatorial argument.
In EFLTL, after every prefix of the input, it is
sufficient to keep a finite set of (non temporal) relations over the
data elements seen in the input, which
summarizes the prefix. This follows from the set
semantics, which appears in Section~\ref{EPLTLRV}, 
where the interpretation of
each subformula, and the auxiliary relations,
at a particular state is a relation, and the interpretation 
of a property a prefix of length $i$ only depends on
these relations. Now, the number of such possible relations
for a fixed number of data elements $N$ is polynomial
in $N$. However, the above property checks whether
a set of values that appear within the relation $v$, between two consecutive appearances
of $r$'s, are the same. For that, one needs to keep a set of
subsets of the elements that appeared in the input so far. 
But the number of such subsets is exponential in $N$. \qed
